---
title: "R Notebook"
output: html_notebook
---

```{r}
# Load the gapminder package
library(gapminder)

# Load the dplyr package
library(dplyr)

# Look at the gapminder dataset
gapminder
```

The filter verb extracts particular observations based on a condition.
Add a filter() line after the pipe (%>%) to extract only the observations from the year 1957. Remember that you use == to compare two values.

```{r}
# Filter the gapminder dataset for the year 1957
gapminder %>%
  filter(year == 1957)
```

You can also use the filter() verb to set two conditions, which could retrieve a single observation.

Just like in the last exercise, you can do this in two lines of code, starting with gapminder %>% and having the filter() on the second line. Keeping one verb on each line helps keep the code readable. Note that each time, you'll put the pipe %>% at the end of the first line (like gapminder %>%); putting the pipe at the beginning of the second line will throw an error.

Filter the gapminder data to retrieve only the observation from China in the year 2002.

```{r}
# Filter for China in 2002
gapminder %>%
  filter(country == "China", year == 2002)
```

You use arrange() to sort observations in ascending or descending order of a particular variable. In this case, you'll sort the dataset based on the lifeExp variable

Sort the gapminder dataset in ascending order of life expectancy (lifeExp).
Sort the gapminder dataset in descending order of life expectancy.

```{r}
# Sort in ascending order of lifeExp
gapminder %>%
  arrange(lifeExp)
  
# Sort in descending order of lifeExp
gapminder %>%
  arrange(desc(lifeExp))
```


You'll often need to use the pipe operator (%>%) to combine multiple dplyr verbs in a row. In this case, you'll combine a filter() with an arrange() to find the highest population countries in a particular year.

Use filter() to extract observations from just the year 1957, then use arrange() to sort in descending order of population (pop)

```{r}
# Filter for the year 1957, then arrange in descending order of population
gapminder %>%
  filter(year == 1957) %>%
  arrange(desc(pop))
```

Suppose we want life expectancy to be measured in months instead of years: you'd have to multiply the existing value by 12. You can use the mutate() verb to change this column, or to create a new column that's calculated this way.

Use mutate() to change the existing lifeExp column, by multiplying it by 12: 12 * lifeExp.
Use mutate() to add a new column, called lifeExpMonths, calculated as 12 * lifeExp.

```{r}
# Use mutate to change lifeExp to be in months
gapminder %>%
  mutate(lifeExp = lifeExp * 12)

# Use mutate to create a new column called lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = lifeExp * 12)
```

In this exercise, you'll combine all three of the verbs you've learned in this chapter, to find the countries with the highest life expectancy, in months, in the year 2007.

In one sequence of pipes on the gapminder dataset:
filter() for observations from the year 2007,
mutate() to create a column lifeExpMonths, calculated as 12 * lifeExp, and
arrange() in descending order of that new column

```{r}
# Filter, mutate, and arrange the gapminder dataset
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12 * lifeExp) %>%
  arrange(desc(lifeExpMonths))
```






