---
title: "R Notebook"
output: html_notebook
---

```{r}
# Load xts
library(xts)
```


xts objects are simple. Think of them as a matrix of observations combined with an index of corresponding dates and times.

The main xts constructor takes a number of arguments, but the two most important are x for the data and order.by for the index. x must be a vector or matrix. order.by is a vector which must be the same length or number of rows as x, be a proper time or date object (very important!), and be in increasing order.

xts also allows you to bind arbitrary key-value attributes to your data. This lets you keep metadata about your object inside your object. To add these at creation, you simply pass additional name = value arguments to the xts() function.

```{r}
# Create the object data using 5 random numbers
data <- rnorm(5)

# Create dates as a Date class object starting from 2016-01-01
dates <- seq(as.Date("2016-01-01"), length = 5, by = "days")

# Use xts() to create smith
smith <- xts(x = data, order.by = dates)

# Create bday (1899-05-08) using a POSIXct date class object
bday <- as.POSIXct("1899-05-08")

# Create hayek and add a new attribute called born
hayek <- xts(x = data, order.by = dates, born = bday)
```

It is best to think of xts objects as normal R matrices, but with special powers. These powers let you manipulate your data as a function of time, as your data is now self-aware of when it exists in time

```{r}
# View the structure of ex_matrix
str(hayek)

# Extract the 3rd observation of the 2nd column of ex_matrix
hayek[3,1]

```

Now that you can create xts objects, your next task is to examine an xts object from the inside.

At the core of both xts and zoo is a simple R matrix with a few additional attributes. The most important of these attributes is the index. The index holds all the information we need for xts to treat our data as a time series.

When working with time series, it will sometimes be necessary to separate your time series into its core data and index attributes for additional analysis and manipulation. The core data is the matrix portion of xts. You can separate this from the xts object using coredata(). The index portion of the xts object is available using the index() function. Note that both of these functions are methods from the zoo class, which xts extends.

```{r}
# Extract the core data of hayek
hayek_core <- coredata(hayek)

# View the class of hayek_core
class(hayek_core)

# Extract the index of hayek
hayek_index <- index(hayek)

# View the class of hayek_index
class(hayek_index)
```


xts objects get their power from the index attribute that holds the time dimension. One major difference between xts and most other time series objects in R is the ability to use any one of various classes that are used to represent time. Whether POSIXct, Date, or some other class, xts will convert this into an internal form to make subsetting as natural to the user as possible.

For now you can simply use date objects to index appropriate rows from your time series. You can think of this as effectively matching the rownames you see in the object. This works as anticipated for time objects because the rownames are really dates!

```{r}
# Create dates
dates <- as.Date("2016-01-01") + 0:4

# Create ts_a
ts_a <- xts(x = 1:5, order.by = dates)

# Create ts_b
ts_b <- xts(x = 1:5, order.by = as.POSIXct(dates))

# Extract the rows of ts_a using the index of ts_b
ts_a[index(ts_b)]

# Extract the rows of ts_b using the index of ts_a
ts_b[index(ts_a)]

```

It is often necessary to convert between classes when working with time series data in R. Conversion can be required for many reasons, but typically you'll be looking to use a function that may not be time series aware or you may want to use a particular aspect of xts with something that doesn't necessarily need to be a full time series.

Luckily, it is quite easy to convert back and forth using the standard as.* style functionality provided in R (for example, as.POSIXct() or as.matrix()).

xts provides methods to convert all of the major objects you are likely to come across. Suitable native R types like matrix, data.frame, and ts are supported, as well as contributed ones such as timeSeries, fts and of course zoo. as.xts() is the workhorse function to do the conversions to xts, and similar functions will provide the reverse behavior.

```{r}
# Convert austres to an xts object called au
au <- as.xts(austres)

# Then convert your xts object (au) into a matrix am
am <- as.matrix(au)

# Inspect the head of am
head(am)

# Convert the original austres into a matrix am2
am2 <- as.matrix(austres)

# Inspect the head of am2
head(am2)

```

You can now convert data to xts using as.xts(). However, in most real world applications you will often need to read raw data from files on disk or the web. This can be challenging without knowing the right commands.

Read in a csv file from disk using the base-R read.csv. After you read the data, the next step is to convert it to xts. Here you will be required to use the xts() constructor as well as deal with converting non-standard dates into something that xts understands.

```{r}
# Create data by reading in file from the Internet
data <- read.csv("https://s3.amazonaws.com/assets.datacamp.com/production/course_1127/datasets/tmp_file.csv")

# Convert data into xts named data_ts
 data_ts <- xts(data, order.by = as.Date(rownames(data), "%m/%d/%Y"))

```






















