---
title: "R Notebook"
output: html_notebook
---


Load the ggplot2 package after the gapminder and dplyr packages.
Filter gapminder for observations from the year 1952, and assign it to a new dataset gapminder_1952 using the assignment operator (<-).

To assign a new dataset, you would do gapminder_1952 <-, then create your filtered dataset with dplyr's filter().

```{r}
# Load the ggplot2 package as well
library(gapminder)
library(dplyr)
library(ggplot2)

# Create gapminder_1952
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
```

Change the scatter plot of gapminder_1952 so that (pop) is on the x-axis and GDP per capita (gdpPercap) is on the y-axis.

```{r}

# Change to put pop on the x-axis and gdpPercap on the y-axis
ggplot(gapminder_1952, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```

Use ggplot2 to create a scatter plot from scratch, to compare each country's population with its life expectancy in the year 1952.

Create a scatter plot of gapminder_1952 with population (pop) is on the x-axis and life expectancy (lifeExp) on the y-axis.

```{r}
# Create a scatter plot with pop on the x-axis and lifeExp on the y-axis
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point()
```

You previously created a scatter plot with population on the x-axis and life expectancy on the y-axis. Since population is spread over several orders of magnitude, with some countries having a much higher population than others, it's a good idea to put the x-axis on a log scale.

Change the existing scatter plot (code provided) to put the x-axis (representing population) on a log scale.


```{r}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

# Change this plot to put the x-axis on a log scale
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() +
  scale_x_log10()
```






