---
title: "R Notebook"
output: html_notebook
---


The white noise (WN) model is a basic time series model. It is also a basis for the more elaborate models we will consider. We will focus on the simplest form of WN, independent and identically distributed data.

The arima.sim() function can be used to simulate data from a variety of time series models. ARIMA is an abbreviation for the autoregressive integrated moving average class of models we will consider throughout this course.

An ARIMA(p, d, q) model has three parts, the autoregressive order p, the order of integration (or differencing) d, and the moving average order q. We will detail each of these parts soon, but for now we note that the ARIMA(0, 0, 0) model, i.e., with all of these components zero, is simply the WN model.

```{r}
# Simulate a WN model with list(order = c(0, 0, 0))
white_noise<- arima.sim(model = list(order = c(0, 0, 0)), n = 100)

# Plot the WN observations
ts.plot(white_noise)
```

```{r}
# Simulate from the WN model with: mean = 100, sd = 10
white_noise_2 <- arima.sim(list(order = c(0, 0, 0)), n = 100, mean = 100, sd = 10)

# Plot your white_noise_2 data
ts.plot(white_noise_2)
```




For a given time series white_noise_2 we can fit the white noise (WN) model using the arima(..., order = c(0, 0, 0)) function. Recall that the WN model is an ARIMA(0,0,0) model. Applying the arima() function returns information or output about the estimated model. For the WN model this includes the estimated mean, labeled intercept, and the estimated variance, labeled sigma^2.

In this exercise, you'll explore the qualities of the WN model. What is the estimated mean? Compare this with the sample mean using the mean() function. What is the estimated variance? Compare this with the sample variance using the var() function.

```{r}
# Fit the WN model to y using the arima command
arima(white_noise_2, order = c(0,0,0))

# Calculate the sample mean and sample variance of y
mean(white_noise_2)
var(white_noise_2)
```







